Channel Shaping Using Beyond Diagonal RIS: Analysis, Optimization, and Enhanced Flexibility

### 2024-11-20 poster/beamer.tex
:lipstick: Update poster

					The capability of \gls{d} and \gls{bd} \gls{ris} to redistribute the singular values of \glsfmtshort{mimo} channels.
					\item \textbf{How does it differ from previous work?}

@@

			\begin{block}{\gls{bd}-\gls{ris} model}
				\setlength{\leftskip}{\leftmargini}
				\setlength{\rightskip}{\leftmargini}
				Consider an $N_\mathrm{T} \times N_\mathrm{S} \times N_\mathrm{R}$ setup with \gls{bd}-\gls{ris} divided into $G$ groups of $L$ elements each.
				Define $N = \min(N_\mathrm{T}, N_\mathrm{R})$, $\bar{N} = \max(N_\mathrm{T}, N_\mathrm{S}, N_\mathrm{R})$, and $\mathbf{H}_{\mathrm{B}/\mathrm{F}}\stackrel{\svd}{=} \mathbf{U}_{\mathrm{B}/\mathrm{F}} \boldsymbol{\Sigma}_{\mathrm{B}/\mathrm{F}} \mathbf{V}_{\mathrm{B}/\mathrm{F}}^\mathsf{H}$.
				\begin{equation*}
					\mathbf{\Theta} = \diag(\mathbf{\Theta}_1,\ldots,\mathbf{\Theta}_G),\quad \mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g = \mathbf{I}_L \ \forall g, \quad \mathbf{H} = \overbrace{\mathbf{H}_\mathrm{D} + \sum\nolimits_g \underbrace{\mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g}}_{\small{\substack{\text{backward-forward:}\\\text{intra-group, multiplicative}}}}}^{\small{\substack{\text{direct-indirect:}\\\text{inter-group, additive}}}}.
				\end{equation*}

@@

				Assume the \gls{bd}-\gls{ris} is symmetric (i.e., $\beta = \pi / 2$) and the direct channel is negligible (i.e., $\sv(e^{\jmath \phi} \mathbf{A}) = \sv(\mathbf{A})$).
				The plot relies on grid search for one channel realization.
				\begin{figure}
					\centering
					\includegraphics[width=0.44\columnwidth]{../assets/simulation/singular_trend.eps}
					% \caption{
					% 	\setlength{\leftskip}{\leftmargini}
					% 	\setlength{\rightskip}{\leftmargini}
					% 	Singular value shaping by \gls{d}-\gls{ris} and symmetric fully-connected \gls{bd}-\gls{ris}.
					% 	The maximum and minimum singular values are marked explicitly on the plot.}
				\end{figure}
				Here, both singular values are manipulated up to $\pm 9\%$ by \gls{d}-\gls{ris} and $\pm 42\%$ by symmetric fully-connected \gls{bd}-\gls{ris}, using 2 and 3 circuit components respectively.
			\end{exampleblock}
		\end{column}

@@

		\begin{column}{\colwidth}
			\vspace{-0.5cm}
			\begin{algo}{Group-wise geodesic optimization}{}
				\setlength{\leftskip}{\leftmargini}
				\setlength{\rightskip}{\leftmargini}
				% A geodesic is a curve representing the shortest path between two points in a Riemannian manifold.
				For \gls{bd}-\gls{ris} optimization problem of the form
				\begin{maxi*}
					{\scriptstyle{\mathbf{\Theta}}}{f(\mathbf{\Theta})}{}{}
					\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I},}{\quad \forall g,}{}
				\end{maxi*}
				% At iteration $r$, the $g$-th group is updated by \textcolor{imperialblue}{geodesic} \gls{rcg}
				we propose a \textcolor{pink!50!pink!75!black}{geodesic} \gls{rcg} algorithm below.
				\vspace{0.5cm}

				\begin{enumerate}\color{black}
					\item \emph{Compute the Euclidean gradient:}
					\begin{equation*}
						\nabla_{\mathrm{E},g}^{(r)} = \frac{\partial f(\mathbf{\Theta}_g^{(r)})}{\partial \mathbf{\Theta}_g^*};
					\end{equation*}
					\item \emph{Translate to the Riemannian gradient evaluated at the identity:}
					\begin{equation*}
						\tilde{\nabla}_{\mathrm{R},g}^{(r)} = \nabla_{\mathrm{E},g}^{(r)} \mathbf{\Theta}_g^{(r)\mathsf{H}} - \mathbf{\Theta}_g^{(r)} {\nabla_{\mathrm{E},g}^{(r)\mathsf{H}}};
					\end{equation*}
					\item \emph{Determine the conjugate direction:}
					\begin{equation*}
						{\mathbf{D}}_g^{(r)} = \tilde{\nabla}_{\mathrm{R},g}^{(r)} + \tilde{\gamma}_g^{(r)} {\mathbf{D}}_g^{(r-1)}, \quad \tilde{\gamma}_g^{(r)} = \frac{\tr\bigl((\tilde{\nabla}_{\mathrm{R},g}^{(r)} - \tilde{\nabla}_{\mathrm{R},g}^{(r-1)}) {\tilde{\nabla}_{\mathrm{R},g}^{(r)\mathsf{H}}}\bigr)}{\tr\bigl(\tilde{\nabla}_{\mathrm{R},g}^{(r-1)} {\tilde{\nabla}_{\mathrm{R},g}^{(r-1)\mathsf{H}}}\bigr)};
					\end{equation*}
					\item \emph{Perform \textcolor{pink!50!pink!75!black}{multiplicative} update along geodesic:}
					\begin{equation*}
						\mathbf{\Theta}_g^{(r+1)} = \mathbf{G}_g^{(r)}(\mu) = \exp(\mu \mathbf{D}_g^{(r)}) \mathbf{\Theta}_g^{(r)},
					\end{equation*}
					where $\mu$ is refinable by the Armijo rule.
				\end{enumerate}
				\vspace{0.5cm}
			\end{algo}

@@

			\begin{exampleblock}{Result 2: Bounds and Pareto frontier of singular values}
				\begin{figure}[!t]
					\centering
					\subfloat[Analytical vs numerical results: $4 \times 128 \times 4, \rank(\mathbf{H}_\mathrm{F}) = 2$]{
						\resizebox{0.492\linewidth}{!}{
							\input{../assets/poster/singular_bound_rank2_sx128.tex}
						}
					}
					\subfloat[Pareto frontier: $2 \times 64 \times 2$]{
						\resizebox{0.462\linewidth}{!}{
							\input{../assets/poster/singular_pareto_sx64.tex}
						}
					}
				\end{figure}
			\end{exampleblock}

			\begin{exampleblock}{Result 3: Channel power gain and achievable rate}
				\begin{figure}[!t]
					\centering
					\subfloat[Channel power gain: $16 \times N_\mathrm{S} \times 16$]{
						\resizebox{0.492\linewidth}{!}{
							\input{../assets/poster/power_sx_txrx16.tex}
						}
					}
					\subfloat[Achievable rate: $N_\mathrm{T} \times 128 \times N_\mathrm{R}$]{
						\resizebox{0.462\linewidth}{!}{
							\input{../assets/poster/rate_txrx.tex}
						}
					}
				\end{figure}

>>>>

					The capability of \gls{d}- and \gls{bd}-\gls{ris} to redistribute the singular values of \glsfmtshort{mimo} channels.
					\item \textbf{How does it differ from previous work?}

@@

			\begin{block}{\gls{bd}-\gls{ris} model}
				\setlength{\leftskip}{\leftmargini}
				\setlength{\rightskip}{\leftmargini}
				Consider an $N_\mathrm{T} \times N_\mathrm{S} \times N_\mathrm{R}$ setup with \gls{bd}-\gls{ris} divided into $G$ groups of $L$ elements each.
				Define $N = \min(N_\mathrm{T}, N_\mathrm{R})$ and $\mathbf{H}_{\mathrm{B}/\mathrm{F}}\stackrel{\svd}{=} \mathbf{U}_{\mathrm{B}/\mathrm{F}} \boldsymbol{\Sigma}_{\mathrm{B}/\mathrm{F}} \mathbf{V}_{\mathrm{B}/\mathrm{F}}^\mathsf{H}$.
				\begin{equation*}
					\mathbf{\Theta} = \diag(\mathbf{\Theta}_1,\ldots,\mathbf{\Theta}_G),\quad \mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g = \mathbf{I}_L \ \forall g, \quad \mathbf{H} = \overbrace{\mathbf{H}_\mathrm{D} + \sum\nolimits_g \underbrace{\mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g}}_{\small{\substack{\text{backward-forward:}\\\text{intra-group, multiplicative}}}}}^{\small{\substack{\text{direct-indirect:}\\\text{inter-group, additive}}}}.
				\end{equation*}

@@

				Assume the \gls{bd}-\gls{ris} is symmetric (i.e., $\beta = \pi / 2$) and the direct channel is negligible (i.e., $\sv(e^{\jmath \phi} \mathbf{A}) = \sv(\mathbf{A})$).
				For one channel realization, we can reveal channel singular values achieved by \gls{d}- and \gls{bd}-\gls{ris} by grid search over $(\theta_1, \theta_2)$ and $(\alpha, \psi)$.
				\begin{figure}
					\centering
					\includegraphics[width=0.41\columnwidth]{../assets/simulation/singular_trend.eps}
					% \caption{
					% 	\setlength{\leftskip}{\leftmargini}
					% 	\setlength{\rightskip}{\leftmargini}
					% 	Singular value shaping by \gls{d}-\gls{ris} and symmetric fully-connected \gls{bd}-\gls{ris}.
					% 	The maximum and minimum singular values are marked explicitly on the plot.}
				\end{figure}
				Here, both singular values are manipulated up to $\pm 9\%$ by \gls{d}-\gls{ris} and $\pm 42\%$ by symmetric fully-connected \gls{bd}-\gls{ris}, using 2 and 3 circuit components respectively.
			\end{exampleblock}
		\end{column}

@@

		\begin{column}{\colwidth}
			\vspace{-0.5cm}
			\begin{algo}{Group-wise geodesic optimization for \gls{bd}-\gls{ris}}{}
				\setlength{\leftskip}{\leftmargini}
				\setlength{\rightskip}{\leftmargini}
				% A geodesic is a curve representing the shortest path between two points in a Riemannian manifold.
				For \gls{bd}-\gls{ris} optimization problem of the form
				\begin{maxi*}
					{\scriptstyle{\mathbf{\Theta}}}{f(\mathbf{\Theta})}{}{}
					\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I},}{\quad \forall g,}{}
				\end{maxi*}
				% At iteration $r$, the $g$-th group is updated by \textcolor{imperialblue}{geodesic} \gls{rcg}
				we propose a \textcolor{pink!50!pink!75!black}{geodesic} \gls{rcg} algorithm below.
				\vspace{0.5cm}

				\begin{enumerate}\color{black}
					\item \emph{Compute the Euclidean gradient:}
					\begin{equation*}
						\nabla_{\mathrm{E},g}^{(r)} = \frac{\partial f(\mathbf{\Theta}_g^{(r)})}{\partial \mathbf{\Theta}_g^*};
					\end{equation*}
					\item \emph{Translate to the Riemannian gradient evaluated at the identity:}
					\begin{equation*}
						\tilde{\nabla}_{\mathrm{R},g}^{(r)} = \nabla_{\mathrm{E},g}^{(r)} \mathbf{\Theta}_g^{(r)\mathsf{H}} - \mathbf{\Theta}_g^{(r)} {\nabla_{\mathrm{E},g}^{(r)\mathsf{H}}};
					\end{equation*}
					\item \emph{Determine the conjugate direction:}
					\begin{equation*}
						{\mathbf{D}}_g^{(r)} = \tilde{\nabla}_{\mathrm{R},g}^{(r)} + \tilde{\gamma}_g^{(r)} {\mathbf{D}}_g^{(r-1)}, \quad \tilde{\gamma}_g^{(r)} = \frac{\tr\bigl((\tilde{\nabla}_{\mathrm{R},g}^{(r)} - \tilde{\nabla}_{\mathrm{R},g}^{(r-1)}) {\tilde{\nabla}_{\mathrm{R},g}^{(r)\mathsf{H}}}\bigr)}{\tr\bigl(\tilde{\nabla}_{\mathrm{R},g}^{(r-1)} {\tilde{\nabla}_{\mathrm{R},g}^{(r-1)\mathsf{H}}}\bigr)};
					\end{equation*}
					\item \emph{Perform \textcolor{pink!50!pink!75!black}{multiplicative} update along geodesic:}
					\begin{equation*}
						\mathbf{\Theta}_g^{(r+1)} = \mathbf{G}_g^{(r)}(\mu) = \exp(\mu \mathbf{D}_g^{(r)}) \mathbf{\Theta}_g^{(r)},
					\end{equation*}
					where $\mu$ is refinable by the Armijo rule. To double the step size, one can simply square the rotation matrix instead of recomputing the matrix exponential.
				\end{enumerate}
				% \vspace{0.5cm}
			\end{algo}

@@

			\begin{exampleblock}{Result 2: Bounds and Pareto frontier of singular values}
				\begin{figure}[!t]
					\centering
					\subfloat[Analytical vs numerical results: $4 \times 128 \times 4, \rank(\mathbf{H}_\mathrm{F}) = 2$]{
						\resizebox{0.485\linewidth}{!}{
							\input{../assets/poster/singular_bound_rank2_sx128.tex}
						}
					}
					\subfloat[Pareto frontier: $2 \times 64 \times 2$]{
						\resizebox{0.454\linewidth}{!}{
							\input{../assets/poster/singular_pareto_sx64.tex}
						}
					}
				\end{figure}
			\end{exampleblock}

			\begin{exampleblock}{Result 3: Channel power gain and achievable rate}
				\begin{figure}[!t]
					\centering
					\subfloat[Channel power gain: $16 \times N_\mathrm{S} \times 16$]{
						\resizebox{0.485\linewidth}{!}{
							\input{../assets/poster/power_sx_txrx16.tex}
						}
					}
					\subfloat[Achievable rate: $N_\mathrm{T} \times 128 \times N_\mathrm{R}$]{
						\resizebox{0.454\linewidth}{!}{
							\input{../assets/poster/rate_txrx.tex}
						}
					}
				\end{figure}
